---
title: "ClaimStat TSW"
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows

---



```{r global, include=FALSE}
library(flexdashboard)
library(data.table)
library(dygraphs)
library(DT)
library(dplyr)
library(plyr)
library(leaflet)
library(plotly)
library(rsconnect)
library(shiny)

#setwd("~/Documents/SD_data/TSW_dashboard_050719")
claims <- read.csv("/usr/local/airflow/poseidon/data/prod/claim_stat_datasd.csv", header=T, na.strings=c(""," ","NA"))

# CHANGE ME FOR PRODUCTION URL!!!!!!!
# claims <- read.csv("http://datasd-dev.s3.amazonaws.com/risk/claims_clean_datasd_v1.csv", header=T, na.strings=c("", " ", "NA"))
claims<-unique(claims)

claims <-data.table(claims)
claims$INCIDENT_DATE<- as.Date(claims$INCIDENT_DATE)
claims$incident_year = format(claims$INCIDENT_DATE, "%Y")

###Create year and month variables
claims$INCIDENT_DATE<- as.Date(claims$INCIDENT_DATE)
claims$incident_year<-format(claims$INCIDENT_DATE, "%Y")
claims$incident_month<-format(claims$INCIDENT_DATE, "%m")


##Create FISCAL YEAR variable
#This variable considers that fiscal years start the first day of July (first day of month 7)
##and run through the last day of June
claims$Fiscal_Year<- ifelse(as.numeric(claims$incident_month)>= 7, (as.numeric(claims$incident_year)+1), claims$incident_year)
claims<-claims[order(-Fiscal_Year),]
claims$Fiscal_Year<- as.numeric(claims$Fiscal_Year)
claims<- subset(claims, Fiscal_Year >=2009)


###Create reserves variable
claims$reserves=claims$INCURRED_TOTAL-claims$PAID_TOTAL


```

Department Loss Experience
=======================================================================

Inputs {.sidebar data-width=200}
-----------------------------------------------------------------------

```{r}


 selectInput("year", label = "Fiscal Year of Incident",
    choices = claims$Fiscal_Year,
    selected=1,
    multiple=FALSE
   )

```

Outputs
-----------------------------------------------------------------------
### Total Payments to Date {.value-box}

```{r}
renderValueBox({
  filtered<-subset(claims, Fiscal_Year == input$year)
total_amount<- round(sum(filtered$PAID_TOTAL, na.rm=TRUE))
total_amount<- paste("$", total_amount, sep="")



  valueBox((prettyNum(total_amount, big.mark= ",")), icon = "fa-dollar", color="orange")
})
```

### Total Claims Count {.value-box}

```{r}


# Emit the user count
renderValueBox({
  filtered<-subset(claims, Fiscal_Year == input$year)
  total_claims<-length(unique(filtered$CLAIM_NUMBER))
  valueBox((prettyNum(total_claims, big.mark= ",")), icon = "fa-receipt")
})

```



Row {.tabset}
-----------------------------------------------------------------------
### Payments to Date {.tabset}
Share of payments by Loss Code: includes all Loss Codes amounting to 1% or more of total payments. For further detail refer to Loss Code Table.



```{r}
renderPlotly({

filtered<-subset(claims, Fiscal_Year == input$year)
options(scipen = 999)

test<-setDT(filtered)[, .(Losses= sum(PAID_TOTAL, na.rm = TRUE), Count=n_distinct(CLAIM_NUMBER)), by = .(CLAIMANT_REFERENCE2_Desc)]
test$percentage<-test$Losses/sum(filtered$PAID_TOTAL, na.rm=TRUE)
test<-test[order(-percentage)]
#select top only
test<- subset(test, percentage>=0.005)
test<-droplevels(test)

test <- data.frame(test, stringsAsFactors = FALSE)
test$CLAIMANT_REFERENCE2_Desc <- factor(test$CLAIMANT_REFERENCE2_Desc, levels = unique(test$CLAIMANT_REFERENCE2_Desc)[order(test$percentage, decreasing = TRUE)])

#Format ##
test$Losses<- format(test$Losses, big.mark=",", scientific=FALSE, digits=0)


  plot_ly(test, x = ~CLAIMANT_REFERENCE2_Desc, y = ~percentage, type = 'bar', name = 'Total Claims', text=~paste('Total Paid: $', Losses,
                      ' </br>  Claims Count: ', Count
                      )) %>%
   layout(margin = list(b= 220), yaxis = list(title = 'Share of Payments by Code'), xaxis = list(title = 'Loss Code', tickangle = -60),barmode = 'group') %>%
  layout(yaxis = list(tickformat = "%"))


  })

```


### Reserves
Share of outstanding reserves by Loss Code: includes all Loss Codes amounting to 1% or more of total outstanding reserves.
*Note that only open claims have outstanding reserves, for incident years with no open claims, no graph will be displayed*

```{r}
renderPlotly({

filtered<-subset(claims, Fiscal_Year == input$year & CLAIMANT_STATUS_DESC=="Open")
options(scipen = 999)

test<-setDT(filtered)[, .(Losses= sum(PAID_TOTAL, na.rm = TRUE), Reserves=sum(reserves, na.rm = TRUE), Count=n_distinct(CLAIM_NUMBER)), by = .(CLAIMANT_REFERENCE2_Desc)]
test$percentage<-test$Reserves/sum(filtered$reserves, na.rm=TRUE)
test<-test[order(-percentage)]

##select top only
test<- subset(test, percentage>=0.005)
test<-droplevels(test)

test <- data.frame(test, stringsAsFactors = FALSE)
test$CLAIMANT_REFERENCE2_Desc <- factor(test$CLAIMANT_REFERENCE2_Desc, levels = unique(test$CLAIMANT_REFERENCE2_Desc)[order(test$percentage, decreasing = TRUE)])

#Format ##
test$Losses<- format(test$Losses, big.mark=",", scientific=FALSE, digits=0)
test$Reserves<- format(test$Reserves, big.mark=",", scientific=FALSE, digits=0)



  plot_ly(test, x = ~CLAIMANT_REFERENCE2_Desc, y = ~percentage, type = 'bar', name = 'Total Claims', text=~paste('Reserves: $', Reserves,
                      ' </br> Open Claims Count: ', Count,
                      '</br> Total Paid: $', Losses)) %>%
   layout(margin = list(b= 220), yaxis = list(title = 'Share of Reserves by Code'), xaxis = list(title = 'Loss Code', tickangle = -60),barmode = 'group') %>%
  layout(yaxis = list(tickformat = "%"))


  })

```

### Loss Code Table
Loss leaders are shown on top of the table: Loss codes are ranked by the share of total payments they represent for the selected FY of incident.
```{r}

renderTable({

##filter data by year selected in sidebar
filtered<-subset(claims, Fiscal_Year == input$year)

##create data table
##############
##all claims
##############
options(scipen = 999)
test<-setDT(filtered)[, .(Total_Losses= sum(PAID_TOTAL, na.rm = TRUE), Count=n_distinct(CLAIM_NUMBER)), by = .(CLAIMANT_REFERENCE2_Desc)]
test$percentage<-(test$Total_Losses/sum(filtered$PAID_TOTAL, na.rm=TRUE))*100

test<-test[order(-percentage)]


##############
## Reserves
##############
filtered3<-subset(filtered, CLAIMANT_STATUS_DESC=="Open")

test2<-setDT(filtered3)[, .(Reserves=sum(reserves, na.rm = TRUE) ), by = .(CLAIMANT_REFERENCE2_Desc)]

####join total with open and closed
final<-join(test, test2, by="CLAIMANT_REFERENCE2_Desc", type="left")

############################
##Format variables
############################
final$percentage<- paste(round(final$percentage,digits=2),"%",sep="")
final$Total_Losses<-prettyNum(final$Total_Losses, big.mark= ",", digits=0)
final$Total_Losses<- paste("$", final$Total_Losses, sep="")


final$Reserves<-prettyNum(final$Reserves, big.mark= ",", digits=0)
final$Reserves<- paste("$", final$Reserves, sep="")

##Remove na Loss codes
final<-final[!is.na(final$CLAIMANT_REFERENCE2_Desc),]

############################
##Create data frame for table
############################

  final %>%
    select("Loss Code" = CLAIMANT_REFERENCE2_Desc, "% of Total Payments" = percentage, "Total Payments to Date"= Total_Losses, "Claims Count"= Count,
                                                     "Reserves"= Reserves) %>%
    as.data.frame()


}, digits = 0)
```






Map
=======================================================================

Inputs {.sidebar data-width=200}
-----------------------------------------------------------------------

```{r}
##Remove NA loss codes
map<- subset(claims, CLAIMANT_REFERENCE2_Desc!= "NA")

#Remove lat-longs clearly outside of SDC
#map<- subset(map, lng>=-117.27 & lng<=-117.11 & lat >=32.567 & lat <=32.88)
#map<- subset(map, lng>=-117.27 & lng<=-117.04 & lat >=32.567 & lat <=32.88)
map<- subset(map, lng>=-117.27 & lng<=-117.04 & lat >=32.567 & lat <=33.06)


#Remove rows with missing lat-longs
map<-map[!is.na(map$lng),]


##Sort data by number of claims per loss code
map<- map %>% group_by(CLAIMANT_REFERENCE2_Desc) %>% mutate(count = n())
map<-map[order(-map$count),]

##Remove rows with generic geocodes (lat long of SD City)
##lat= 32.715738, lng= -117.1610838
#(32.715738, -117.1610838)

map <- map[!(map$lat == "32.715738" & map$lng=="-117.1610838" ),]



 selectInput("losscode", label = "Loss Code",
    choices = map$CLAIMANT_REFERENCE2_Desc,
    selected=1,
    multiple=FALSE
   )



```

Outputs
-----------------------------------------------------------------------
### TSW claims by location {data-width=350}



```{r}
renderLeaflet({
  filtered<-subset(map, CLAIMANT_REFERENCE2_Desc == input$losscode)


leaflet(filtered) %>% addTiles() %>% addMarkers(
  label= ~paste('Claim Number:', CLAIM_NUMBER, '*   Claim Status:', CLAIMANT_STATUS_DESC, '*    Total Paid: $', PAID_TOTAL),
  labelOptions= labelOptions( style= list( "font-size"= "15px", "font-style"= "bold")),
  clusterOptions = markerClusterOptions()
)

})
```

Time Trends
=======================================================================

```{r}


renderPlotly({

##change this if I include a sidebar filter: FILTER BY COUNT OR AMOUNT, and by open/closed
test<-setDT(claims)[, .(PAID_TOTAL= sum(PAID_TOTAL, na.rm = TRUE), PAID_BIPD= sum(PAID_BI + PAID_PD, na.rm = TRUE), PAID_EXP=sum(PAID_EXPENSE),Count=n_distinct(CLAIM_NUMBER)), by = .(Fiscal_Year)]

x <- test$Fiscal_Year
y <- test$PAID_TOTAL
trace_1<- test$PAID_BIPD
trace_2<- test$PAID_EXP

p <- plot_ly(x = ~x, y = ~y, mode = 'lines')%>%
add_trace(y = ~y, name = 'Total Payments', mode = 'lines+markers')%>%
add_trace(y = ~trace_1, name = 'Bodily Injury and Property Damage', mode = 'lines+markers')%>%
add_trace(y = ~trace_2, name = 'Expenses', mode = 'lines+markers') %>%
  layout(title = "Payment History by Fiscal Year of Incident",
         xaxis = list(title = "Fiscal Year of Incident"),
         yaxis = list (title = "Total Amount (USD)"))
})




```






Claim Level Detail
=======================================================================
### Claims Search Tool {data-width=350}

```{r}
claims$INCIDENT_DATE<-as.Date(claims$INCIDENT_DATE)
claims_sub<- claims[, c("INCIDENT_DATE","Fiscal_Year", "CLAIM_NUMBER","CLAIMANT_STATUS_DESC","PAID_TOTAL", "CLAIMANT_REFERENCE2_Desc", "INCIDENT_DESC", "ORGANIZATION_DESC")]
claims_sub$Fiscal_Year<-as.factor(claims_sub$Fiscal_Year)
claims_sub<-claims_sub[!is.na(claims_sub$CLAIMANT_REFERENCE2_Desc),]
claims_sub<-claims_sub[order(claims_sub$CLAIMANT_STATUS_DESC),]

datatable(claims_sub, rownames = FALSE, filter = 'top', colnames = c( 'Incident Date', 'Fiscal Year','Claim Number','Claim Status','Total Payment to Date', 'Loss Code', 'Incident Description', 'Organization Description'), options = list(
  pageLength = 10, autoWidth = TRUE, columnDefs = list(list(width = '100px', targets = "_all"))
))
```


About
=======================================================================
### ClaimStat Tool guide {data-width=350}

For descriptions of each of the components of ClaimStat, as well as a Loss Code Dictionary, please visit the ClaimStat Tool Guide by going to the link below

https://docs.google.com/spreadsheets/d/18WmU6Hifz4OaaupHqwKLhBJFZTGuwH3tEt6CaWiP1cs/edit?usp=sharing
