---
title: "ClaimStat PD"
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows

---



```{r global, include=FALSE}
library(flexdashboard)
library(data.table)
library(dygraphs)
library(DT)
library(dplyr)
library(plyr)
library(leaflet)
library(plotly)
library(rsconnect)
library(shiny)
library(purrr)
library(colorspace)


claims <- read.csv("http://seshat.datasd.org/risk/claims_clean_datasd_pd.csv", header=T, na.strings=c(""," ","NA"))

# claims <- read.csv("http://datasd-dev.s3.amazonaws.com/risk/claims_clean_datasd_v1.csv", header=T, na.strings=c("", " ", "NA"))
#claims <- read.csv("PD_test.csv", header=T, na.strings=c("", " ", "NA")) #TEMP FILE FOR TESTING JOB
claims<-unique(claims)

claims <-data.table(claims)
claims$INCIDENT_DATE<- as.Date(claims$INCIDENT_DATE)
claims$incident_year = format(claims$INCIDENT_DATE, "%Y")

###Create year and month variables
claims$INCIDENT_DATE<- as.Date(claims$INCIDENT_DATE)
claims$incident_year<-format(claims$INCIDENT_DATE, "%Y")
claims$incident_month<-format(claims$INCIDENT_DATE, "%m")


##Create FISCAL YEAR variable
#This variable considers that fiscal years start the first day of July (first day of month 7)
##and run through the last day of June
claims$Fiscal_Year<- ifelse(as.numeric(claims$incident_month)>= 7, (as.numeric(claims$incident_year)+1), claims$incident_year)
claims<-claims[order(-Fiscal_Year),]
claims$Fiscal_Year<- as.numeric(claims$Fiscal_Year)
claims<- subset(claims, Fiscal_Year >=2009)


###Create reserves variable
claims$reserves=claims$INCURRED_TOTAL-claims$PAID_TOTAL


```

Department Loss Experience
=======================================================================

Inputs {.sidebar data-width=200}
-----------------------------------------------------------------------

```{r}


 selectInput("year", label = "Fiscal Year of Incident",
    choices = claims$Fiscal_Year,
    selected=1,
    multiple=FALSE
   )

```

Outputs
-----------------------------------------------------------------------
### Total Payments to Date {.value-box}

```{r}
renderValueBox({
  filtered<-subset(claims, Fiscal_Year == input$year)
total_amount<- round(sum(filtered$PAID_TOTAL, na.rm=TRUE))
total_amount<- paste("$", total_amount, sep="")



  valueBox((prettyNum(total_amount, big.mark= ",")), icon = "fa-dollar", color="orange")
})
```

### Total Claims Count {.value-box}

```{r}


# Emit the user count
renderValueBox({
  filtered<-subset(claims, Fiscal_Year == input$year)
  total_claims<-length(unique(filtered$CLAIM_NUMBER))
  valueBox((prettyNum(total_claims, big.mark= ",")), icon = "fa-receipt")
})

```



Row {.tabset}
-----------------------------------------------------------------------
### Payments to Date {.tabset}
Share of payments by Loss Code: includes all Loss Codes amounting to 1% or more of total payments. For further detail refer to Loss Code Table.



```{r}
renderPlotly({

filtered<-subset(claims, Fiscal_Year == input$year)
options(scipen = 999)

test<-setDT(filtered)[, .(Losses= sum(PAID_TOTAL, na.rm = TRUE), Count=n_distinct(CLAIM_NUMBER)), by = .(CLAIMANT_REFERENCE2_Desc)]
test$percentage<-test$Losses/sum(filtered$PAID_TOTAL, na.rm=TRUE)
test<-test[order(-percentage)]
#select top only
test<- subset(test, percentage>=0.005)
test<-droplevels(test)

test <- data.frame(test, stringsAsFactors = FALSE)
test$CLAIMANT_REFERENCE2_Desc <- factor(test$CLAIMANT_REFERENCE2_Desc, levels = unique(test$CLAIMANT_REFERENCE2_Desc)[order(test$percentage, decreasing = TRUE)])

#Format ##
test$Losses<- format(test$Losses, big.mark=",", scientific=FALSE, digits=0)


  plot_ly(test, x = ~CLAIMANT_REFERENCE2_Desc, y = ~percentage, type = 'bar', name = 'Total Claims', text=~paste('Total Paid: $', Losses,
                      ' </br>  Claims Count: ', Count
                      )) %>%
   layout(margin = list(b= 220), yaxis = list(title = 'Share of Payments by Code'), xaxis = list(title = 'Loss Code', tickangle = -60),barmode = 'group') %>%
  layout(yaxis = list(tickformat = "%"))


  })

```


### Reserves
Share of outstanding reserves by Loss Code: includes all Loss Codes amounting to 1% or more of total outstanding reserves.
*Note that only open claims have outstanding reserves, for incident years with no open claims, no graph will be displayed*

```{r}
renderPlotly({

filtered<-subset(claims, Fiscal_Year == input$year & CLAIMANT_STATUS_DESC=="Open")
options(scipen = 999)

test<-setDT(filtered)[, .(Losses= sum(PAID_TOTAL, na.rm = TRUE), Reserves=sum(reserves, na.rm = TRUE), Count=n_distinct(CLAIM_NUMBER)), by = .(CLAIMANT_REFERENCE2_Desc)]
test$percentage<-test$Reserves/sum(filtered$reserves, na.rm=TRUE)
test<-test[order(-percentage)]

##select top only
test<- subset(test, percentage>=0.005)
test<-droplevels(test)

test <- data.frame(test, stringsAsFactors = FALSE)
test$CLAIMANT_REFERENCE2_Desc <- factor(test$CLAIMANT_REFERENCE2_Desc, levels = unique(test$CLAIMANT_REFERENCE2_Desc)[order(test$percentage, decreasing = TRUE)])

#Format ##
test$Losses<- format(test$Losses, big.mark=",", scientific=FALSE, digits=0)
test$Reserves<- format(test$Reserves, big.mark=",", scientific=FALSE, digits=0)



  plot_ly(test, x = ~CLAIMANT_REFERENCE2_Desc, y = ~percentage, type = 'bar', name = 'Total Claims', text=~paste('Reserves: $', Reserves,
                      ' </br> Open Claims Count: ', Count,
                      '</br> Total Paid: $', Losses)) %>%
   layout(margin = list(b= 220), yaxis = list(title = 'Share of Reserves by Code'), xaxis = list(title = 'Loss Code', tickangle = -60),barmode = 'group') %>%
  layout(yaxis = list(tickformat = "%"))


  })

```

### Loss Code Table
Loss leaders are shown on top of the table: Loss codes are ranked by the share of total payments they represent for the selected FY of incident.
```{r}

renderTable({

##filter data by year selected in sidebar
filtered<-subset(claims, Fiscal_Year == input$year)

##create data table
##############
##all claims
##############
options(scipen = 999)
test<-setDT(filtered)[, .(Total_Losses= sum(PAID_TOTAL, na.rm = TRUE), Count=n_distinct(CLAIM_NUMBER)), by = .(CLAIMANT_REFERENCE2_Desc)]
test$percentage<-(test$Total_Losses/sum(filtered$PAID_TOTAL, na.rm=TRUE))*100

test<-test[order(-percentage)]


##############
## Reserves
##############
filtered3<-subset(filtered, CLAIMANT_STATUS_DESC=="Open")

test2<-setDT(filtered3)[, .(Reserves=sum(reserves, na.rm = TRUE) ), by = .(CLAIMANT_REFERENCE2_Desc)]

####join total with open and closed
final<-join(test, test2, by="CLAIMANT_REFERENCE2_Desc", type="left")

############################
##Format variables
############################
final$percentage<- paste(round(final$percentage,digits=2),"%",sep="")
final$Total_Losses<-prettyNum(final$Total_Losses, big.mark= ",", digits=0)
final$Total_Losses<- paste("$", final$Total_Losses, sep="")


final$Reserves<-prettyNum(final$Reserves, big.mark= ",", digits=0)
final$Reserves<- paste("$", final$Reserves, sep="")

##Remove na Loss codes
final<-final[!is.na(final$CLAIMANT_REFERENCE2_Desc),]

############################
##Create data frame for table
############################

  final %>%
    select("Loss Code" = CLAIMANT_REFERENCE2_Desc, "% of Total Payments" = percentage, "Total Payments to Date"= Total_Losses, "Claims Count"= Count,
                                                     "Reserves"= Reserves) %>%
    as.data.frame()


}, digits = 0)
```






Map
=======================================================================


```{r}
##Remove NA loss codes
map<- subset(claims, CLAIMANT_REFERENCE2_Desc== "0260-City Vehicle Accident")

#Remove lat-longs clearly outside of SDC
#map<- subset(map, lng>=-117.27 & lng<=-117.11 & lat >=32.567 & lat <=32.88)
map<- subset(map, lng>=-117.27 & lng<=-117.04 & lat >=32.567 & lat <=32.88)


#Remove rows with missing lat-longs
map<-map[!is.na(map$lng),]



##Remove rows with generic geocodes (lat long of SD City)
##lat= 32.715738, lng= -117.1610838
#(32.715738, -117.1610838)

map <- map[!(map$lat == "32.715738" & map$lng=="-117.1610838" ),]





```

-----------------------------------------------------------------------
### City Vehicle Accident PD claims by location {data-width=350} 

```{r}
renderLeaflet({

leaflet(map) %>% addTiles() %>% addMarkers(
  label= ~paste('Claim Number:', CLAIM_NUMBER, '*   Claim Status:', CLAIMANT_STATUS_DESC, '*    Total Paid: $', PAID_TOTAL),
  labelOptions= labelOptions( style= list( "font-size"= "15px", "font-style"= "bold")),
  clusterOptions = markerClusterOptions()
)

})
```



Time Trends (top loss codes) 
=======================================================================
-----------------------------------------------------------------------
## Frequency

```{r}

renderPlotly({

test<-setDT(claims)[, .(PAID_TOTAL= sum(PAID_TOTAL, na.rm = TRUE),Count=n_distinct(CLAIM_NUMBER)), 
                by = .(Fiscal_Year, CLAIMANT_REFERENCE2_Desc)]

top_5_count<-setDT(claims)[, .(Count=n_distinct(CLAIM_NUMBER, na.rm = TRUE)), 
                         by = .(CLAIMANT_REFERENCE2_Desc)]

top_5_count<- top_5_count[order(-Count),]
top_5_count <- top_5_count[1:5,]
top_5_count$top<-1
top_5_count<-top_5_count[,c(1,3)]
top_5_count<-left_join(top_5_count, test)
top_5_count<-select(top_5_count, CLAIMANT_REFERENCE2_Desc,Fiscal_Year, Count)

top_5_count<-subset(top_5_count, Fiscal_Year<2020) #TEMPORARILY, JUST FOR PROTOTYPE:


top_5_count <- reshape(top_5_count, 
                     timevar = "CLAIMANT_REFERENCE2_Desc",
                     idvar = c("Fiscal_Year"),
                     direction = "wide")

top_5_count<-top_5_count %>%
  set_names(~stringr::str_replace_all(., "Count.", " ")
  )


x <- top_5_count$Fiscal_Year
top_1<-unlist(c(top_5_count[,2]))
top_2<-unlist(c(top_5_count[,3]))
top_3<-unlist(c(top_5_count[,4]))
top_4<-unlist(c(top_5_count[,5]))
top_5<-unlist(c(top_5_count[,6]))


freqs <- plot_ly(x = ~x, y = ~top_1, mode = 'lines')%>%
  add_trace(y = ~top_1, name = names(top_5_count[1,2]), mode = 'lines+markers',marker = list(color = 'grey'), line = list(color = 'grey'))%>%
  add_trace(y = ~top_2, name = names(top_5_count[1,3]), mode = 'lines+markers')%>%
  add_trace(y = ~top_3, name = names(top_5_count[1,4]), mode = 'lines+markers' ,marker = list(color = 'saddlebrown'), line = list(color = 'saddlebrown'))%>%
  add_trace(y = ~top_4, name = names(top_5_count[1,5]), mode = 'lines+markers')%>%
  add_trace(y = ~top_5, name = names(top_5_count[1,6]), mode = 'lines+markers')%>%
  layout(
         xaxis = list(title = "Fiscal Year of Incident", dtick=1),
         yaxis = list (title = "Frequency (claims count)"))






})

```
 
-----------------------------------------------------------------------
## Severity

```{r}


renderPlotly({

test<-setDT(claims)[, .(PAID_TOTAL= sum(PAID_TOTAL, na.rm = TRUE),Count=n_distinct(CLAIM_NUMBER)), 
                by = .(Fiscal_Year, CLAIMANT_REFERENCE2_Desc)]

top_5_usd<-setDT(claims)[, .(PAID_TOTAL= sum(PAID_TOTAL, na.rm = TRUE)), 
                    by = .(CLAIMANT_REFERENCE2_Desc)]

top_5_usd<- top_5_usd[order(-PAID_TOTAL),]
top_5_usd <- top_5_usd[1:5,]
top_5_usd$top<-1
top_5_usd<-top_5_usd[,c(1,3)]
top_5_usd<-left_join(top_5_usd, test)
top_5_usd<-select(top_5_usd, CLAIMANT_REFERENCE2_Desc,Fiscal_Year, PAID_TOTAL)

top_5_usd<-subset(top_5_usd, Fiscal_Year<2020) #TEMPORARILY, JUST FOR PROTOTYPE:


top_5_usd <- reshape(top_5_usd, 
             timevar = "CLAIMANT_REFERENCE2_Desc",
             idvar = c("Fiscal_Year"),
             direction = "wide")

top_5_usd<-top_5_usd %>%
  set_names(~stringr::str_replace_all(., "PAID_TOTAL.", " ")
  )


x <- top_5_usd$Fiscal_Year
top_1<-unlist(c(top_5_usd[,2]))
top_2<-unlist(c(top_5_usd[,3]))
top_3<-unlist(c(top_5_usd[,4]))
top_4<-unlist(c(top_5_usd[,5]))
top_5<-unlist(c(top_5_usd[,6]))


figu <- plot_ly(top_5_usd, x = ~x, y = ~top_1, type = 'bar', name = names(top_5_usd[1,2]))
figu <- figu %>% add_trace(y = ~top_2, name = names(top_5_usd[1,3]))
figu <- figu %>% add_trace(y = ~top_3, name = names(top_5_usd[1,4]))
figu <- figu %>% add_trace(y = ~top_4, name = names(top_5_usd[1,5]) , marker = list(color = 'darkolivegreen'))
figu <- figu %>% add_trace(y = ~top_5, name = names(top_5_usd[1,6]))
figu <- figu %>% layout(yaxis = list(title = 'Severity (claims cost)'), barmode = 'group')
figu <- figu %>% layout(xaxis = list(title = 'Fiscal Year of Incident', dtick=1), barmode = 'group', 
                        autotick= F, dtick = 1, ticktext=x)


figu



})




```
 















Claim Level Detail
=======================================================================
### Claims Search Tool {data-width=350}

```{r}
claims$INCIDENT_DATE<-as.Date(claims$INCIDENT_DATE)
claims_sub<- claims[, c("INCIDENT_DATE","Fiscal_Year", "CLAIM_NUMBER","CLAIMANT_STATUS_DESC","PAID_TOTAL", "CLAIMANT_REFERENCE2_Desc", "INCIDENT_DESC", "ORGANIZATION_DESC")]
claims_sub$Fiscal_Year<-as.factor(claims_sub$Fiscal_Year)
claims_sub<-claims_sub[!is.na(claims_sub$CLAIMANT_REFERENCE2_Desc),]
claims_sub<-claims_sub[order(claims_sub$CLAIMANT_STATUS_DESC),]

datatable(claims_sub, rownames = FALSE, filter = 'top', colnames = c( 'Incident Date', 'Fiscal Year','Claim Number','Claim Status','Total Payment to Date', 'Loss Code', 'Incident Description', 'Organization Description'), options = list(
  pageLength = 10, autoWidth = TRUE, columnDefs = list(list(width = '100px', targets = "_all"))
))
```


About
=======================================================================
### ClaimStat Tool guide {data-width=350}

For descriptions of each of the components of ClaimStat, as well as a Loss Code Dictionary, please visit the ClaimStat Tool Guide by going to the link below

https://docs.google.com/spreadsheets/d/112b_rjybntIVE-_sKqJcJwjvdTc47FAf92pUKO9x78M/edit?usp=sharing

